\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2018-08-08 paper/ms.tex
New replicate.py and correspondingly updated results

\section{Postprocessing}\label{sec:postprocessing}
The output files themselves are typically not immediately useful.
The goal of running
nested sampling is usually to obtain posterior samples and the marginal likelihood $Z$,
whereas \code{sample.txt} only contains samples from the mixture of constrained
priors. Additional
post-processing is required. This can be achieved by running the following
\proglang{Python} function\footnote{Alternatively, the file
\code{showresults.py} in the example directory runs this function, and then
calls the code in \code{display.py} to create a further plot shown in
Figure~\ref{fig:fig3} on the right.}:
%
\begin{CodeChunk}
\begin{CodeInput}
>>> import dnest4
>>> dnest4.postprocess()
\end{CodeInput}
\end{CodeChunk}
%
This produces the three diagnostic plots in Figures~\ref{fig:fig1}
and~\ref{fig:fig3} on the left, along with the following output:
%
\begin{CodeChunk}
\begin{CodeOutput}
log(Z) = -175.49577676640516
Information = 15.17486407892028 nats.
Effective sample size = 1126.5312086470944
\end{CodeOutput}
\end{CodeChunk}
%
These are the natural log of the marginal likelihood, the
information
\begin{align*}
\mathcal{H} &= \int p(\theta|D, M)
\ln\left[\frac{p(\theta | D, M)}{p(\theta | M)}\right] \, d\theta,
\end{align*}
which quantifies the degree to which $D$ restricted the
range of possible $\theta$ values,
and the effective sample size,
or number of saved particles with significant posterior weight.
The \code{postprocess} function also saves a file,
\code{posterior\_sample.txt}, containing posterior samples (one per row).
Unfortunately, it is harder to compute justified error bars on $\ln(Z)$
in DNS than it is in standard nested sampling.
%The \code{postprocess} function does include some options
%for calculating error bars, but they are not used by default,
%and we do not describe them here.
%Hence, the zero-sized error bars on $\ln(Z)$ and $\mathcal{H}$ should
%be ignored.

@@

    vec[0] = l1.cdf(vec[0]);
    vec[1] = l2.cdf(vec[1]);
    vec[2] = l3.cdf(vec[2]);
}
\end{CodeInput}
\end{CodeChunk}
%
\subsection{Mixture model results}
The data and the posterior mean fit are shown
in Figure~\ref{fig:galaxies} on the left and the posterior distribution for
$N$ is shown in Figure~\ref{fig:galaxies} on the right.
The marginal likelihood was $\ln(Z) = -232.2$, and
the information (measured in nats) was $\mathcal{H} = 29.5$.

@@

    return logL;
}
\end{CodeInput}
\end{CodeChunk}
%
\subsection{ABC example results}
For ABC applications, there is an alternate \code{postprocess} function
called \code{postprocess_abc}.
This generates posterior samples using the $\epsilon$ value corresponding
to level $\approx 3/4\times$~(maximum number of levels).
%
\begin{CodeChunk}
\begin{CodeInput}
dnest4.postprocess_abc()
\end{CodeInput}
\end{CodeChunk}
%
To try different values of $\epsilon$, you can set the argument
\code{threshold\_fraction} to a value other than 0.8. For example
%
\begin{CodeChunk}
\begin{CodeInput}
dnest4.postprocess_abc(threshold_fraction = 0.6)
\end{CodeInput}
\end{CodeChunk}
%
In the example directory there is a dataset generated from
a standard normal distribution. The inference should
result in a joint posterior for $\mu$ and $\sigma$ with
significant density near $(\mu=0, \sigma=1)$.
We show posterior samples in Figure~\ref{fig:abc_results}.
\begin{figure}[t!]
\centering
\includegraphics[width=0.6\textwidth, trim = 0 5 0 25]{abc_results.pdf}
\caption{Posterior samples for $\mu$ and $\sigma$ for the
ABC demonstration.\label{fig:abc_results}}
\end{figure}

>>>>

\section{Postprocessing}\label{sec:postprocessing}
The output files themselves are typically not immediately useful.
The goal of running
nested sampling is usually to obtain posterior samples and the marginal likelihood $Z$,
whereas \code{sample.txt} only contains samples from the mixture of constrained
priors. Additional
post-processing is required. This can be achieved by running the following
\proglang{Python} function\footnote{Alternatively, the file
\code{showresults.py} in the example directory runs this function, and then
calls the code in \code{display.py} to create a further plot shown in
Figure~\ref{fig:fig3} on the right.}:
%
\begin{CodeChunk}
\begin{CodeInput}
>>> import dnest4
>>> dnest4.postprocess()
\end{CodeInput}
\end{CodeChunk}
%
This produces the three diagnostic plots in Figures~\ref{fig:fig1}
and~\ref{fig:fig3} on the left, along with the following output:
%
\begin{CodeChunk}
\begin{CodeOutput}
log(Z) = -175.47872682863076
Information = 15.166340875399044 nats.
Effective sample size = 1251.3576752091165
\end{CodeOutput}
\end{CodeChunk}
%
These are the natural log of the marginal likelihood, the
information
\begin{align*}
\mathcal{H} &= \int p(\theta|D, M)
\ln\left[\frac{p(\theta | D, M)}{p(\theta | M)}\right] \, d\theta,
\end{align*}
which quantifies the degree to which $D$ restricted the
range of possible $\theta$ values,
and the effective sample size,
or number of saved particles with significant posterior weight.
The \code{postprocess} function also saves a file,
\code{posterior\_sample.txt}, containing posterior samples (one per row).
Unfortunately, it is harder to compute justified error bars on $\ln(Z)$
in DNS than it is in standard nested sampling.
%The \code{postprocess} function does include some options
%for calculating error bars, but they are not used by default,
%and we do not describe them here.
%Hence, the zero-sized error bars on $\ln(Z)$ and $\mathcal{H}$ should
%be ignored.

@@

    vec[0] = l1.cdf(vec[0]);
    vec[1] = l2.cdf(vec[1]);
    vec[2] = l3.cdf(vec[2]);
}
\end{CodeInput}
\end{CodeChunk}
%
\subsection{Mixture model results}
The data and the posterior mean fit are shown
in Figure~\ref{fig:galaxies} on the left and the posterior distribution for
$N$ is shown in Figure~\ref{fig:galaxies} on the right.
The marginal likelihood was $\ln(Z) = -232.1$, and
the information was $\mathcal{H} = 29.5$ nats.

@@

    return logL;
}
\end{CodeInput}
\end{CodeChunk}
%
\subsection{ABC example results}
For ABC applications, there is an alternate \code{postprocess} function
called \code{postprocess_abc}.
This generates posterior samples using the $\epsilon$ value corresponding
to level $\approx 0.8\times$~(maximum number of levels).
%
\begin{CodeChunk}
\begin{CodeInput}
dnest4.postprocess_abc()
\end{CodeInput}
\end{CodeChunk}
%
To try different values of $\epsilon$, you can set the argument
\code{threshold\_fraction} to a value other than 0.8. For example
%
\begin{CodeChunk}
\begin{CodeInput}
dnest4.postprocess_abc(threshold_fraction = 0.6)
\end{CodeInput}
\end{CodeChunk}
%
In the example directory there is a dataset generated from
a standard normal distribution. The inference should
result in a joint posterior for $\mu$ and $\sigma$ with
significant density near $(\mu=0, \sigma=1)$.
We show posterior samples in Figure~\ref{fig:abc_results}.
\begin{figure}[t!]
\centering
\includegraphics[width=0.6\textwidth, trim = 0 5 0 25]{abc_results.pdf}
\caption{Posterior samples for $\mu$ and $\sigma$ for the
ABC demonstration.\label{fig:abc_results}}
\end{figure}
